We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-01-24 paper/paper.tex
edit

In this paper we develop and demostrate a method to quantify the
asphericity of the satellite distribution in the Local Group.
The method uses as a reference point the spherically randomized data
of the system under study.
We first measure the width and axis ratios for the satellite
distribution of interest. Then, we measure the same quantities for the spherical
randomizations, to finally renormalize the initial results to the mean
value and standard deviation from the randomized samples.

We do not look for planes  or coherent kinematical structures in the
interest of keeping our study simple and robust.
In observations we limit our analysis to the 11 brightest satellites 
in the Milky Way and Andromeda. 
For numerical data we use the Illustris-1, Illustris-1-DM and the
ELVIS simulations.

We find that the deviation from aesphericity in the LG is only
expected in $1\pm3$ pairs out of a sample of $2000$ isolated pairs. 
This places the LG as a $3\sigma$ outlier. 
The weight to explain this atypical result is not distributed equally
between the MW and M31. While M31 presents a fully typical asphericity
in the expectations from LCDM, the MW shows aspherical deviations in
plane width and the major-to-minor axis ratio highly atypical in the
framework of LCDM. 
We estimate that with the M31 $XX$ out of $2000$ pairs show normalized
characteristics larget r equal than M31, while this number drops to 
$XX$ for the MW.

>>>>

In this paper we develop and demostrate a method to quantify the
asphericity of the satellite distribution in the Local Group.
We build up on the suggestions made in \cite{XX}.
We do not look for planes  or coherent kinematical structures in the
interest of keeping the method straightforward and robust.

The method uses as a reference point the spherically randomized data
of the system under study.
We first measure the width and axis ratios for the satellite
distribution of interest. 
Then, we measure the same quantities for spherically
randomized data. 
Finally, renormalize the initial results to the mean value and
standard deviation from the randomized samples.

The normalized quantities for different systems are well reproduced by
a multivariate gaussian distribution. 
We fit the parameters of this gaussian to the results from 
LG samples coming from three different simulations: ELVIS, Illustris1
and Illustris-DarkMatter.
We finally compare the observational results against the distributions
derived from the simulations.

We find that in the best case (Illustris1-DarkMatter) the deviation from aesphericity in the observed LG is only
expected in $3\pm2$ pairs out of a sample of $10^4$ isolated pairs. 
This places the LG as a $3\sigma$ outlier. 
The weight to explain this atypical result is not distributed equally
between the MW and M31. While M31 presents a fully typical asphericity
in the expectations from LCDM, the MW shows aspherical deviations in
plane width and the major-to-minor axis ratio highly atypical in the
framework of LCDM. 
We estimate that with the M31 $XX$ out of $2000$ pairs show normalized
characteristics larget r equal than M31, while this number drops to 
$XX$ for the MW.
